---
title: "Bayesian-models_genetic-diversity_resilience"
authors: "MW, SCGM, MK"
date: "2026-05-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This script contains the code used to fit hierarchical models in the study "Genetic diversity underpins forest ecosystem resilience".


## Load packages

```{r cars}
library(dplyr)
library(ggplot2)
library(posterior)
library(brms)
library(tidyverse)
#library(loo)
library(posterior)
library(bayesplot)
```

## Load & prepare data

```{r}
# Load the data (adjust separator if needed: "\t" for tab, "," for CSV)
#setwd("/media/TrueNAS_analize/resilience_scripts")
data23sps <- read.table("data23sps_213GCUs_Bayes.txt", header = TRUE, row.names=NULL)
data23sps$species<-as.factor(data23sps$species)
data23sps$GENSID<-as.factor(data23sps$GENSID)

# Standardize Hs by species
data23sps$stHs <- unlist(by(data23sps, data23sps$species, FUN = function(x) scale(x$Hs)))

```


The models aim to assess the relationship between genetic diversity metrics (Hs, π) and ecosystem resilience as measured by restoration rates (λAC1, λvar).


# Model 1 

Because species were genotyped using different technologies, which could have biased genetic diversity estimates (SPET vs. SNP arrays; see Pinosio et al. 2026), the model fit for all 23 species was based on genetic diversity estimate (Hs) standardised at the species level.

```{r pressure, echo=FALSE}
# ========================================================================================================================
# MODEL 1.1: B = stHs, A = λAC1
# ========================================================================================================================

# Create an output folder
model_name <- "fit_m1.1_stHs_10k_iter"
out_dir <- paste0(model_name, "_files")
dir.create(out_dir, showWarnings = FALSE)

# Fit the model
fit_m1.1_stHs_10k_iter <- brm(
  formula = lambda_AC1_19Y ~ stHs + (1|GENSID),
  data = data23sps,
  family = gaussian(),
  chains = 4,
  iter = 10000,
  warmup = 1000,
  seed = 12345,
  control = list(adapt_delta = 0.99, max_treedepth = 15)
)

# Summary
print(summary(fit_m1.1_stHs_10k_iter))
write.csv(as.data.frame(fixef(fit_m1.1_stHs_10k_iter)),
          file.path(out_dir, paste0(model_name, "_fixef.csv")))
hypothesis(fit_m1.1_stHs_10k_iter, "stHs < 0")
fixef(fit_m1.1_stHs_10k_iter)
pp_check(fit_m1.1_stHs_10k_iter)

#  Convergence diagnostics
summ <- posterior::summarise_draws(as_draws_df(fit_m1.1_stHs_10k_iter))
print(summ %>% select(variable, rhat, ess_bulk, ess_tail))
write.csv(summ, file.path(out_dir, paste0(model_name, "_convergence_summary.csv")))

cat("\nParameters with Rhat > 1.01:\n")
print(summ %>% filter(rhat > 1.01) %>% select(variable, rhat))

cat("\nParameters with bulk ESS < 400:\n")
print(summ %>% filter(ess_bulk < 400) %>% select(variable, ess_bulk))

np <- nuts_params(fit_m1.1_stHs_10k_iter)
n_div <- sum(np$Parameter == "divergent__" & np$Value == 1)
cat("\nNumber of divergences:", n_div, "\n")

# Trace plots
pdf(file.path(out_dir, paste0(model_name, "_traceplots.pdf")), width = 10, height = 8)
plot(fit_m1.1_stHs_10k_iter)
dev.off()

# Posterior predictive check
pdf(file.path(out_dir, paste0(model_name, "_ppcheck.pdf")), width = 8, height = 6)
pp_check(fit_m1.1_stHs_10k_iter)
dev.off()

# Model evaluation
loo_m1.1_stHs_10k_iter <- brms::loo(fit_m1.1_stHs_10k_iter)
print(loo_m1.1_stHs_10k_iter)
saveRDS(loo_m1.1_stHs_10k_iter, file.path(out_dir, paste0(model_name, "_loo.rds")))

# Conditional effects
pdf(file.path(out_dir, paste0(model_name, "_conditional_effects.pdf")), width = 8, height = 6)
plot(conditional_effects(fit_m1.1_stHs_10k_iter), ask = FALSE)
dev.off()

# Save model
saveRDS(fit_m1.1_stHs_10k_iter, file.path(out_dir, paste0(model_name, ".rds")))


# ----------------------------------------------------------------------
# Visualization - example for m1.1 - lambda_AC1_19Y, stHs, 23 species
# ----------------------------------------------------------------------

# Extract the model-implied relationship 
ce <- conditional_effects(fit_m1.1_stHs_10k_iter, effects = "stHs")
ce_df <- as.data.frame(ce$stHs)

# Plot: raw data points + model-based fitted line & CI ribbon
ggplot() +
    geom_point(data = data23sps, aes(x = stHs, y = lambda_AC1_19Y),
               color = "#2E8B57", size = 2.5, alpha = 0.7) +
    geom_ribbon(data = ce_df, aes(x = stHs, ymin = lower__, ymax = upper__),
                fill = "#2E8B57", alpha = 0.25) +
    geom_line(data = ce_df, aes(x = stHs, y = estimate__),
              color = "#2E8B57", linewidth = 1.2) +
    labs(x = expression(italic(stH[s])), y = expression(lambda[AC1])) +
    theme_bw(base_size = 14) +
    theme(
      panel.grid = element_blank(),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
      axis.title = element_text(size = 16),
      axis.text = element_text(color = "black"),
      aspect.ratio = 1
  )




# ========================================================================================================================
# MODEL 1.2: B = stHs, A = λvar
# ========================================================================================================================

# Create an output folder
model_name <- "fit_m1.2_stHs_10k_iter"
out_dir <- paste0(model_name, "_files")
dir.create(out_dir, showWarnings = FALSE)

# Fit the model
fit_m1.2_stHs_10k_iter <- brm(
  formula = lambda_variance_19Y ~ stHs + (1|GENSID),
  data = data23sps,
  family = gaussian(),
  chains = 4,
  iter = 10000,
  warmup = 1000,
  seed = 12345,
  control = list(adapt_delta = 0.99, max_treedepth = 15)
)

# Summary
print(summary(fit_m1.2_stHs_10k_iter))
write.csv(as.data.frame(fixef(fit_m1.2_stHs_10k_iter)),
          file.path(out_dir, paste0(model_name, "_fixef.csv")))
hypothesis(fit_m1.2_stHs_10k_iter, "stHs < 0")
fixef(fit_m1.2_stHs_10k_iter)
pp_check(fit_m1.2_stHs_10k_iter)

#  Convergence diagnostics
summ <- posterior::summarise_draws(as_draws_df(fit_m1.2_stHs_10k_iter))
print(summ %>% select(variable, rhat, ess_bulk, ess_tail))
write.csv(summ, file.path(out_dir, paste0(model_name, "_convergence_summary.csv")))

cat("\nParameters with Rhat > 1.01:\n")
print(summ %>% filter(rhat > 1.01) %>% select(variable, rhat))

cat("\nParameters with bulk ESS < 400:\n")
print(summ %>% filter(ess_bulk < 400) %>% select(variable, ess_bulk))

np <- nuts_params(fit_m1.2_stHs_10k_iter)
n_div <- sum(np$Parameter == "divergent__" & np$Value == 1)
cat("\nNumber of divergences:", n_div, "\n")

# Trace plots
pdf(file.path(out_dir, paste0(model_name, "_traceplots.pdf")), width = 10, height = 8)
plot(fit_m1.2_stHs_10k_iter)
dev.off()

# Posterior predictive check
pdf(file.path(out_dir, paste0(model_name, "_ppcheck.pdf")), width = 8, height = 6)
pp_check(fit_m1.2_stHs_10k_iter)
dev.off()

# Model evaluation
loo_m1.2_stHs_10k_iter <- brms::loo(fit_m1.2_stHs_10k_iter)
print(loo_m1.2_stHs_10k_iter)
saveRDS(loo_m1.2_stHs_10k_iter, file.path(out_dir, paste0(model_name, "_loo.rds")))

# Conditional effects
pdf(file.path(out_dir, paste0(model_name, "_conditional_effects.pdf")), width = 8, height = 6)
plot(conditional_effects(fit_m1.2_stHs_10k_iter), ask = FALSE)
dev.off()

# Save model
saveRDS(fit_m1.2_stHs_10k_iter, file.path(out_dir, paste0(model_name, ".rds")))
```


# Model 2

The second model was similar to the first one, but only used the 17 species genotyped with Single primer enrichment technology (SPET). Therefore, it did not standardise Hs at the species level and excluded *Abies nebrodensis*, which was represented by the only existing population, while accounting for tree species (1|sp).

```{r}
# Exclude A.nebrodensis GCU and species that were not genotyped with SPET (keep only species genotyped with the same technology)
data17sps <- subset(
  data23sps,
  !species %in% c("Pinus_pinaster", "Pinus_pinea", "Pinus_sylvestris", "Populus_alba", "Populus_nigra", "Abies_nebrodensis")
)
```

Fit model 2
```{r}
# ========================================================================================================================
# MODEL 2.1 - example for B = Hs, A = λAC1
# ========================================================================================================================

# Create an output folder
model_name <- "fit_m2.1_Hs_10k_iter"
out_dir <- paste0(model_name, "_files")
dir.create(out_dir, showWarnings = FALSE)

# Fit the model
fit_m2.1_Hs_10k_iter <- brm(
  formula = lambda_xt_19Y ~ Hs + (1|species) + (1|GENSID),
  data = data17sps,
  family = gaussian(),
  chains = 4,
  iter = 10000,
  warmup = 1000,
  seed = 12345,
  control = list(adapt_delta = 0.99, max_treedepth = 15)
)

# Summary
print(summary(fit_m2.1_Hs_10k_iter))
write.csv(as.data.frame(fixef(fit_m2.1_Hs_10k_iter)),
          file.path(out_dir, paste0(model_name, "_fixef.csv")))
hypothesis(fit_m2.1_Hs_10k_iter, "Hs < 0")
fixef(fit_m2.1_Hs_10k_iter)
pp_check(fit_m2.1_Hs_10k_iter)

#  Convergence diagnostics
summ <- posterior::summarise_draws(as_draws_df(fit_m2.1_Hs_10k_iter))
print(summ %>% select(variable, rhat, ess_bulk, ess_tail))
print(summ)
write.csv(summ, file.path(out_dir, paste0(model_name, "_convergence_summary.csv")))

cat("\nParameters with Rhat > 1.01:\n")
print(summ %>% filter(rhat > 1.01) %>% select(variable, rhat))

cat("\nParameters with bulk ESS < 400:\n")
print(summ %>% filter(ess_bulk < 400) %>% select(variable, ess_bulk))

np <- nuts_params(fit_m2.1_Hs_10k_iter)
n_div <- sum(np$Parameter == "divergent__" & np$Value == 1)
cat("\nNumber of divergences:", n_div, "\n")

# Trace plots
pdf(file.path(out_dir, paste0(model_name, "_traceplots.pdf")), width = 10, height = 8)
plot(fit_m2.1_Hs_10k_iter)
dev.off()

# Posterior predictive check
pdf(file.path(out_dir, paste0(model_name, "_ppcheck.pdf")), width = 8, height = 6)
pp_check(fit_m2.1_Hs_10k_iter)
dev.off()

# Model evaluation
loo_m2.1_Hs_10k_iter <- brms::loo(fit_m2.1_Hs_10k_iter)
print(loo_m2.1_Hs_10k_iter)
saveRDS(loo_m2.1_Hs_10k_iter, file.path(out_dir, paste0(model_name, "_loo.rds")))

# Conditional effects
pdf(file.path(out_dir, paste0(model_name, "_conditional_effects.pdf")), width = 8, height = 6)
plot(conditional_effects(fit_m2.1_Hs_10k_iter), ask = FALSE)
dev.off()

# Save model
saveRDS(fit_m2.1_Hs_10k_iter, file.path(out_dir, paste0(model_name, ".rds")))

# ----------------------------------------------------------------------
# Visualization - example for m2.1 - lambda_AC1_19Y, Hs, 17 species
# ----------------------------------------------------------------------

# Extract the model-implied relationship 
ce <- conditional_effects(fit_m2.1_Hs_10k_iter, effects = "Hs")
ce_df <- as.data.frame(ce$Hs)

# Plot: raw data points + model-based fitted line & CI ribbon
ggplot() +
  geom_point(data = data17sps, aes(x = Hs, y = lambda_AC1_19Y),
             color = "#2E8B57", size = 2.5, alpha = 0.7,) +
  geom_ribbon(data = ce_df, aes(x = Hs, ymin = lower__, ymax = upper__),
              fill = "#2E8B57", alpha = 0.25) +
  geom_line(data = ce_df, aes(x = Hs, y = estimate__),
            color = "#2E8B57", linewidth = 1.2) +
  labs(x = expression(italic(H[s])), y = expression(lambda[AC1])) +
  theme_bw(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
    axis.title = element_text(size = 16),
    axis.text = element_text(color = "black"),
    aspect.ratio = 1
  )




# ========================================================================================================================
# MODEL 2.2 - example for B = Hs, A = λvar
# ========================================================================================================================

# Create an output folder
model_name <- "fit_m2.2_Hs_10k_iter"
out_dir <- paste0(model_name, "_files")
dir.create(out_dir, showWarnings = FALSE)

# Fit the model
fit_m2.2_Hs_10k_iter <- brm(
  formula = lambda_variance_19Y ~ Hs + (1|species) + (1|GENSID),
  data = data17sps,
  family = gaussian(),
  chains = 4,
  iter = 10000,
  warmup = 1000,
  seed = 12345,
  control = list(adapt_delta = 0.99, max_treedepth = 15)
)

# Summary
print(summary(fit_m2.2_Hs_10k_iter))
write.csv(as.data.frame(fixef(fit_m2.2_Hs_10k_iter)),
          file.path(out_dir, paste0(model_name, "_fixef.csv")))
hypothesis(fit_m2.2_Hs_10k_iter, "Hs < 0")
fixef(fit_m2.2_Hs_10k_iter)
pp_check(fit_m2.2_Hs_10k_iter)

#  Convergence diagnostics
summ <- posterior::summarise_draws(as_draws_df(fit_m2.2_Hs_10k_iter))
print(summ %>% select(variable, rhat, ess_bulk, ess_tail))
write.csv(summ, file.path(out_dir, paste0(model_name, "_convergence_summary.csv")))

cat("\nParameters with Rhat > 1.01:\n")
print(summ %>% filter(rhat > 1.01) %>% select(variable, rhat))

cat("\nParameters with bulk ESS < 400:\n")
print(summ %>% filter(ess_bulk < 400) %>% select(variable, ess_bulk))

np <- nuts_params(fit_m2.2_Hs_10k_iter)
n_div <- sum(np$Parameter == "divergent__" & np$Value == 1)
cat("\nNumber of divergences:", n_div, "\n")

# Trace plots
pdf(file.path(out_dir, paste0(model_name, "_traceplots.pdf")), width = 10, height = 8)
plot(fit_m2.2_Hs_10k_iter)
dev.off()

# Posterior predictive check
pdf(file.path(out_dir, paste0(model_name, "_ppcheck.pdf")), width = 8, height = 6)
pp_check(fit_m2.2_Hs_10k_iter)
dev.off()

# Model evaluation
loo_m2.2_Hs_10k_iter <- brms::loo(fit_m2.2_Hs_10k_iter)
print(loo_m2.2_Hs_10k_iter)
saveRDS(loo_m2.2_Hs_10k_iter, file.path(out_dir, paste0(model_name, "_loo.rds")))

# Conditional effects
pdf(file.path(out_dir, paste0(model_name, "_conditional_effects.pdf")), width = 8, height = 6)
plot(conditional_effects(fit_m2.2_Hs_10k_iter), ask = FALSE)
dev.off()

# Save model
saveRDS(fit_m2.2_Hs_10k_iter, file.path(out_dir, paste0(model_name, ".rds")))

```


# Model 3
We extend model 2 by including a genetic diversity × species interaction term (B × sp) to assess the species-specific component of the B-A relationship, 
```{r}
# ========================================================================================================================
# MODEL 3 - example for B = Hs, A = λAC1
# ========================================================================================================================

# Create an output folder
model_name <- "fit_m3.1_Hs_10k_iter"
out_dir <- paste0(model_name, "_files")
dir.create(out_dir, showWarnings = FALSE)

#Fit the model
fit_m3.1_Hs_10k_iter <- brm(
  formula = lambda_AC1_19Y ~ Hs * species + (1|GENSID),
  data = data17sps,
  family = gaussian(),
  chains = 4,
  iter = 10000,
  warmup = 1000,
  seed = 12345,
  control = list(adapt_delta = 0.99, max_treedepth = 25)
)

# Summary
print(summary(fit_m3.1_Hs_10k_iter))
write.csv(as.data.frame(fixef(fit_m3.1_Hs_10k_iter)),
          file.path(out_dir, paste0(model_name, "_fixef.csv")))
hypothesis(fit_m3.1_Hs_10k_iter, "Hs < 0")

# Species-specific slopes
post <- as_draws_df(fit_m3.1_Hs_10k_iter)
tree_sp <- levels(data17sps$species)
ref_species <- tree_sp[1]

# Reference species slope
ref_slopes <- tibble(
  tree_sp = ref_species,
  slope = post$b_Hs
)

# All other slopes
slopes <- post %>%
  mutate(global_slope = b_Hs) %>%
  select(global_slope, starts_with("b_Hs:species")) %>%
  pivot_longer(-global_slope,
               names_to = "term",
               values_to = "adjustment") %>%
  mutate(
    tree_sp = str_replace(term, "b_Hs:species", ""),
    slope = global_slope + adjustment
  )

# Put together and summary
slopes <- bind_rows(ref_slopes, slopes)

summary_slopes <- slopes %>%
  group_by(tree_sp) %>%
  summarise(
    mean = mean(slope),
    l95 = quantile(slope, 0.05),
    u95 = quantile(slope, 0.95),
    p_negative = mean(slope < 0),
    p_positive = mean(slope > 0),
  ) %>%
  arrange(mean)
write.csv(summary_slopes, "summary_slopes.csv")

# Forest plot
p_forest <- ggplot(summary_slopes, aes(x = reorder(tree_sp, mean), y = mean)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = l95, ymax = u95), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  coord_flip() +
  labs(x = "Species", y = "Species-specific slopes (β)")

pdf(file.path(out_dir, paste0(model_name, "_forest_plot.pdf")), width = 9, height = 6)
print(p_forest)
dev.off()

# Forest plot excluding Q. pubescens
plot_data <- summary_slopes %>% filter(tree_sp != "Quercus_pubescens")
p_forest_no_pub <- ggplot(plot_data, aes(x = reorder(tree_sp, mean), y = mean)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = l95, ymax = u95), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  coord_flip() +
  labs(x = "Species", y = "Species-specific slopes (β)")

pdf(file.path(out_dir, paste0(model_name, "_forest_plot_no_pubescens.pdf")), width = 9, height = 6)
print(p_forest_no_pub)
dev.off()

#  Convergence diagnostics
summ <- posterior::summarise_draws(as_draws_df(fit_m3.1_Hs_10k_iter))
print(summ %>% select(variable, rhat, ess_bulk, ess_tail))
write.csv(summ, file.path(out_dir, paste0(model_name, "_convergence_summary.csv")))

cat("\nParameters with Rhat > 1.01:\n")
print(summ %>% filter(rhat > 1.01) %>% select(variable, rhat))

cat("\nParameters with bulk ESS < 400:\n")
print(summ %>% filter(ess_bulk < 400) %>% select(variable, ess_bulk))

np <- nuts_params(fit_m3.1_Hs_10k_iter)
n_div <- sum(np$Parameter == "divergent__" & np$Value == 1)
cat("\nNumber of divergences:", n_div, "\n")

# Trace plots
pdf(file.path(out_dir, paste0(model_name, "_traceplots.pdf")), width = 10, height = 8)
plot(fit_m3.1_Hs_10k_iter)
dev.off()

# Posterior predictive check
pdf(file.path(out_dir, paste0(model_name, "_ppcheck.pdf")), width = 8, height = 6)
pp_check(fit_m3.1_Hs_10k_iter)
dev.off()

# Model evaluation
loo_m3.1_Hs_10k_iter <- brms::loo(fit_m3.1_Hs_10k_iter)
print(loo_m3.1_Hs_10k_iter)
saveRDS(loo_m3.1_Hs_10k_iter, file.path(out_dir, paste0(model_name, "_loo.rds")))

# Conditional effects
pdf(file.path(out_dir, paste0(model_name, "_conditional_effects.pdf")), width = 8, height = 6)
plot(conditional_effects(fit_m3.1_Hs_10k_iter), ask = FALSE)
dev.off()

# Save model
saveRDS(fit_m3.1_Hs_10k_iter, file.path(out_dir, paste0(model_name, ".rds")))



# ========================================================================================================================
# MODEL 3.2 - example for B = Hs, A = λvar
# ========================================================================================================================

# Create an output folder
model_name <- "fit_m3.2_Hs_10k_iter"
out_dir <- paste0(model_name, "_files")
dir.create(out_dir, showWarnings = FALSE)

#Fit the model
fit_m3.2_Hs_10k_iter <- brm(
  formula = lambda_variance_19Y ~ Hs * species + (1|GENSID),
  data = data17sps,
  family = gaussian(),
  chains = 4,
  iter = 10000,
  warmup = 1000,
  seed = 12345,
  control = list(adapt_delta = 0.99, max_treedepth = 25)
)

# Summary
print(summary(fit_m3.2_Hs_10k_iter))
write.csv(as.data.frame(fixef(fit_m3.2_Hs_10k_iter)),
          file.path(out_dir, paste0(model_name, "_fixef.csv")))
hypothesis(fit_m3.2_Hs_10k_iter, "Hs < 0")

# Species-specific slopes
post <- as_draws_df(fit_m3.2_Hs_10k_iter)
tree_sp <- levels(data17sps$species)   
ref_species <- tree_sp[1]

# Reference species slope
ref_slopes <- tibble(
  tree_sp = ref_species,
  slope = post$b_Hs
)

# All other slopes
slopes <- post %>%
  mutate(global_slope = b_Hs) %>%
  select(global_slope, starts_with("b_Hs:species")) %>%
  pivot_longer(-global_slope,
               names_to = "term",
               values_to = "adjustment") %>%
  mutate(
    tree_sp = str_replace(term, "b_Hs:species", ""),
    slope = global_slope + adjustment
  )

# Put together and summary
slopes <- bind_rows(ref_slopes, slopes)
summary_slopes <- slopes %>%
  group_by(tree_sp) %>%
  summarise(
    mean = mean(slope),
    l95 = quantile(slope, 0.05),
    u95 = quantile(slope, 0.95),
    p_negative = mean(slope < 0),
    p_positive = mean(slope > 0),
  ) %>%
  arrange(mean)
write.csv(summary_slopes, "summary_slopes.csv")


# Forest plot
p_forest <- ggplot(summary_slopes, aes(x = reorder(tree_sp, mean), y = mean)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = l95, ymax = u95), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  coord_flip() +
  labs(x = "Species", y = "Species-specific slopes (β)")

pdf(file.path(out_dir, paste0(model_name, "_forest_plot.pdf")), width = 9, height = 6)
print(p_forest)
dev.off()

# Forest plot excluding Q. pubescens
plot_data <- summary_slopes %>% filter(tree_sp != "Quercus_pubescens")
p_forest_no_pub <- ggplot(plot_data, aes(x = reorder(tree_sp, mean), y = mean)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = l95, ymax = u95), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  coord_flip() +
  labs(x = "Species", y = "Species-specific slopes (β)")

pdf(file.path(out_dir, paste0(model_name, "_forest_plot_no_pubescens.pdf")), width = 9, height = 6)
print(p_forest_no_pub)
dev.off()



#  Convergence diagnostics
summ <- posterior::summarise_draws(as_draws_df(fit_m3.2_Hs_10k_iter))
print(summ %>% select(variable, rhat, ess_bulk, ess_tail))
write.csv(summ, file.path(out_dir, paste0(model_name, "_convergence_summary.csv")))

cat("\nParameters with Rhat > 1.01:\n")
print(summ %>% filter(rhat > 1.01) %>% select(variable, rhat))

cat("\nParameters with bulk ESS < 400:\n")
print(summ %>% filter(ess_bulk < 400) %>% select(variable, ess_bulk))

np <- nuts_params(fit_m3.2_Hs_10k_iter)
n_div <- sum(np$Parameter == "divergent__" & np$Value == 1)
cat("\nNumber of divergences:", n_div, "\n")

# Trace plots
pdf(file.path(out_dir, paste0(model_name, "_traceplots.pdf")), width = 10, height = 8)
plot(fit_m3.2_Hs_10k_iter)
dev.off()

# Posterior predictive check
pdf(file.path(out_dir, paste0(model_name, "_ppcheck.pdf")), width = 8, height = 6)
pp_check(fit_m3.2_Hs_10k_iter)
dev.off()

# Model evaluation
loo_m3.2_Hs_10k_iter <- brms::loo(fit_m3.2_Hs_10k_iter)
print(loo_m3.2_Hs_10k_iter)
saveRDS(loo_m3.2_Hs_10k_iter, file.path(out_dir, paste0(model_name, "_loo.rds")))

# Conditional effects
pdf(file.path(out_dir, paste0(model_name, "_conditional_effects.pdf")), width = 8, height = 6)
plot(conditional_effects(fit_m3.2_Hs_10k_iter), ask = FALSE)
dev.off()

# Save model
saveRDS(fit_m3.2_Hs_10k_iter, file.path(out_dir, paste0(model_name, ".rds")))

```

